Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-06-08 fib2.tex
more editing; shrink fig 2; add another diagram

We consider a 
2-dimensional system with
point-like excitations or quasiparticles known as anyons.
These anyons show exotic statistics under exchange.
The worldlines of these exchanges form \emph{braids}
in (2+1)-dimensions
which act unitarily on (the Hilbert space of) degrees of freedom known as the \emph{fusion space.}
To access this space we define
the \emph{observables} of the system: these
are regions bounded by circles (Wilson loops)
and measure the total charge of the region.
%The algebraic properties of these processes
These dynamics and observables obey algebraic rules
given by a unitary modular tensor category~\cite{Wang2010b} (UMTC).

@@

The state of a collection of anyons
can be described by the observables of the system.
An important difference between abelian and non-abelian
theories is this: the outcome of measurements in an abelian theory
are determined by particle content only.
That is, the fusion space is one-dimensional (a phase).
This is no longer the case in a non-abelian theory.
An example is Fibonacci: $\tau\times\tau=\tau+\vac$
\emph{picture here?}
%specify a state requires nesting multiple wilson lines.

Non-intersecting loops commute (why?), and so
a basis for the state can be built from a maximal
%set of non-intersecting loops.
set of disjoint and nested loops.
\emph{picture?}
%Observables measure the charge within a
%region bounded by a directed closed loop.
For three anyons, we relate two
possible ways of nesting loops %these measurements
via $F$-moves:
\begin{align}
\label{e:fmove}
\includegraphics[]{pic-fmove.pdf}
\end{align}
Here we have shown particles with charges $a$, $b$, and $c$
as well as total charge $d$.
The lhs shows a state where the anyons with charge $a$ and $b$ are
observed to have total charge $i$.
The rhs involves a superposition (summation)
over charge states $j$ of the anyons with charge $b$ and $c$.
The green line serves to linearly
order the anyons, and track
deformations in the underlying space 
(the history of braiding processes).
It also denotes the direction along
which $F$-moves occur.

%Here we show how braids effect observables:
%Here we show the $R$-move, corresponding to a (clockwise) braiding processes:
A clockwise braiding process is represented by an $R$-move:
\begin{align}
\label{e:rmove}
\includegraphics[]{pic-rmove.pdf}
\end{align}
%where $a$, $b$, and $c$ represent particle types, and $F$-moves,
%representing the associativity rule for fusion processes, 
This is meant to show a \emph{half-twist} of the space
inside the observable with value $c$
holding two anyons (punctures) with charges $a$ and $b$.
We are free to draw loops around charges,
and in general braiding will twist these loops in different
ways.
\emph{picture?}

These pictures, abstracted from the quantum context,
are called \emph{curve diagrams} and are used in
the study of braid groups \cite{Dehornoy2002}.
In general, such a curve diagram is 
a continuous embedding of a line that passes through 
each puncture once. 
Remarkably, any two such linear orderings are related by
braiding (half-twists).

The $R$ and $F$-moves obey composition of diagrams.
\emph{picture?}
Fusion is the ``reverse'' of composition: replace the
interior of a disc by a single charge.
\emph{picture?}

%The dynamics and history of a system of anyons can be conveniently represented by a graphical calculus, which can be regarded as a representation of the world-lines of the particles. These diagrams can be manipulated by $R$-moves, corresponding to braiding processes:
%	\begin{align}
%	\label{e:rmove}
%		\begin{tikzpicture}[baseline=0.5cm,scale=0.3]		
%			\draw(0,0)--(0,1);
%			\draw plot [smooth, tension=1] coordinates { (0,1) (0.75,1.8) (-0.6,2.45) (-1,3)};
%			\draw[color=white, line width=7] plot [smooth, tension=1] coordinates { (0,1) (-0.75,1.8) (0.6,2.45) (1,3)};
%			\draw plot [smooth, tension=1] coordinates { (0,1) (-0.75,1.8) (0.6,2.45) (1,3)};
%			\node at (0,0) [anchor=west] {{\scriptsize $c$}};
%			\node at (-1,3) [anchor=east] {{\scriptsize $b$}};
%			\node at (1,3) [anchor=west] {{\scriptsize $a$}};
%		\end{tikzpicture}
%		&\;\;=\;\;R^{ab}_c\cdot\;
%		\begin{tikzpicture}[baseline=0.5cm,scale=0.3]		
%			\draw(0,0)--(0,1);
%			\draw(0,1)--(-1,2)--(-1,3);
%			\draw(0,1)--(1,2)--(1,3);
%			\node at (0,0) [anchor=west] {{\scriptsize $c$}};
%			\node at (-1,3) [anchor=east] {{\scriptsize $b$}};
%			\node at (1,3) [anchor=west] {{\scriptsize $a$}};
%		\end{tikzpicture}
%	\end{align}
%where $a$, $b$, and $c$ represent particle types, or $F$-moves, representing the associativity rule for fusion processes, 
%	\begin{align}
%		\label{e:fmove}
%		\begin{tikzpicture}[baseline=0.6cm,scale=0.3]
%			\draw(0,0)--(0,1)--(-2,3);
%			\draw(0,3)--(-1,2);
%			\draw(2,3)--(0,1);
%			\node at (-0.45,1.6) [anchor=north east] {{\scriptsize $i$}};
%			\node at (0,0) [anchor=west] {{\scriptsize $d$}};
%			\node at (2,3) [anchor=east] {{\scriptsize $c$}};
%			\node at (0,3) [anchor=east] {{\scriptsize $b$}};
%			\node at (-2,3) [anchor=east] {{\scriptsize $a$}};
%		\end{tikzpicture}
%		&\;\;=\;\;\sum_j[F^{abc}_d]_{ij}\cdot\;
%		\begin{tikzpicture}[baseline=0.6cm,scale=0.3]		
%			\draw(0,0)--(0,1)--(-2,3);
%			\draw(0,3)--(1,2);
%			\draw(2,3)--(0,1);
%			\node at (0.42,1.63) [anchor=north west] {{\scriptsize $j$}};
%			\node at (0,0) [anchor=east] {{\scriptsize $d$}};
%			\node at (2,3) [anchor=east] {{\scriptsize $c$}};
%			\node at (0,3) [anchor=east] {{\scriptsize $b$}};
%			\node at (-2,3) [anchor=east] {{\scriptsize $a$}};
%		\end{tikzpicture}
%	\end{align}

%These manipulations encode the data of the unitary modular tensor 
%category, and can be used to represent the state of 
%a system of anyons as a linear combination of fusion trees.

For the Fibonacci anyons, the non-trivial $R$ and $F$ moves are 
\begin{equation*}
	R_{\vac}^{\tau\tau} = \e^{\frac{-4\pi i}{5}} 
	\ \ , \ \
	R_\tau^{\tau\tau}= \e^{\frac{3\pi i}{5}} 
	\ \ , \ \
	F_{\tau}^{\tau\tau\tau} = \begin{pmatrix}\varphi^{-1}&\varphi^{-\frac{1}{2}}\\\varphi^{-\frac{1}{2}}&-\varphi^{-1}\end{pmatrix} \,,
\end{equation*}
where the matrix is given in a basis labelled $(\vac,\tau)$.

@@

In order to keep track of the state of the system we maintain a set of disjoint curves, 
one for each connected component with trivial total charge.
The application of a pair-creation noise process corresponds to the 
addition of an extra piece of curve to the lattice supporting
two new anyons.
%along with two preferred points on it denoting the anyon locations. 
Following the noise processes, our simulation must %proceed to 
measure the total charge within each tile; the results of 
these measurements will form the error syndrome. 
This requires joining (arbitrarily) any curves that intersect that tile, 
and then performing $R$ and $F$-moves on the resulting
curve so that 
all anyons within the tile have been 
localized within a contiguous region of the curve. 
(Heisenberg picture...)
The observable corresponding to the tile is then
represented in the state ...
%The total charge is then (probabilistically) read off the 
%relevant section of the fusion tree. 
An example of this 
procedure for a simple noise process is shown in \Fref{f:syndrome}.
%and the corresponding braid moves are shown in \Fref{f:syndrome}(e).
The combinatorial derivation of the algorithm and data structures used for 
this procedure will be discussed elsewhere.
%Braiding processes that encircle 
%a non-trivial loop of our (toric) manifold can also be 
%treated in an analogous way, following e.g.~\cite{Pfeifer2012}.
%\cggb{This paragraph and the next still needs another look.}

\begin{figure}[t!]
\begin{center}
	\includegraphics[width=0.8\columnwidth]{pic-syndrome.pdf} %\\[10pt]
%	(e)\ \ 
%	\begin{tikzpicture}[baseline=0.7cm,scale=0.3, xscale=0.9]
%			\begin{scope}[xshift=2cm,yshift=2cm]		
%				\draw(1,0)--(0,1)[dashed];
%				\draw(0,1)--(-1,2)--(-1,3);
%				\draw(0,1)--(1,2)--(1,3);
%				\node at (-1,3) [anchor=west] {{\scriptsize $1$}};
%				\node at (1,3) [anchor=west] {{\scriptsize $2$}};
%			\end{scope}
%			%			
%			\begin{scope}[xshift=4cm]
%				\draw(0,0)--(0,1)[dashed];
%				\draw(0,1)--(-1,2)--(1,4)--(1,5);
%				\draw(0,1)--(3,4)--(3,5);
%				\node at (1,5) [anchor=west] {{\scriptsize $3$}};
%				\node at (3,5) [anchor=west] {{\scriptsize $4$}};
%			\end{scope}
%		\end{tikzpicture}
%		\hspace{-3mm}$\longrightarrow$\hspace{1mm}
%		\begin{tikzpicture}[baseline=1.5cm,scale=0.3, xscale=1]
%			\begin{scope}[xshift=2cm,yshift=2cm]		
%				\draw(1,0)--(0,1)[dashed];
%				\draw(0,1)--(-1,2)--(-1,7);
%				\draw(0,1)--(1,2)--(1,5);
%				\node at (-1,3) [anchor=west] {{\scriptsize $1$}};
%				\node at (1,3) [anchor=west] {{\scriptsize $2$}};
%				\node at (-1,7) [anchor=west] {{\scriptsize $1$}};
%				\node at (1,7) [anchor=west] {{\scriptsize $4$}};
%			\end{scope}
%			%			
%			\begin{scope}[xshift=4cm]
%				\draw(0,0)--(0,1)[dashed];
%				\draw(0,1)--(-1,2)--(1,4)--(1,5);
%				\draw(0,1)--(3,4)--(3,5);
%				\draw(3,7)--(3,9);
%				\node at (1,5) [anchor=west] {{\scriptsize $3$}};
%				\node at (3,5) [anchor=west] {{\scriptsize $4$}};
%				\node at (1,9) [anchor=west] {{\scriptsize $2$}};
%				\node at (3,9) [anchor=west] {{\scriptsize $3$}};
%			\end{scope}
%			%
%			\begin{scope}[xshift=6cm,yshift=2cm]
%				\draw plot [smooth, tension=.6] coordinates { (1,3) (0.75,3.8) (-0.6,4.45) (-1,5)};
%				\draw[color=white, line width=7] plot [smooth, tension=1] coordinates { (-1,3) (-0.75,3.8) (0.6,4.45) (1,5)};
%				\draw plot [smooth, tension=.6] coordinates { (-1,3) (-0.75,3.8) (0.6,4.45) (1,5)};
%			\end{scope}
%			\begin{scope}[xshift=4cm,yshift=4cm]
%				\draw plot [smooth, tension=.6] coordinates { (1,3) (0.75,3.8) (-0.6,4.45) (-1,5)};
%				\draw[color=white, line width=7] plot [smooth, tension=1] coordinates { (-1,3) (-0.75,3.8) (0.6,4.45) (1,5)};
%				\draw plot [smooth, tension=.6] coordinates { (-1,3) (-0.75,3.8) (0.6,4.45) (1,5)};
%			\end{scope}
%		\end{tikzpicture}
%		$\xrightarrow{R, F}$
%		\begin{tikzpicture}[baseline=0.2cm,scale=0.3,xscale=0.9]		
%			\draw(2,-1)--(0,1)[];
%			\draw(0,1)--(-1,2)--(-1,3);
%			\draw(0,1)--(1,2)--(1,3);
%			\node at (-1,3) [anchor=west] {{\scriptsize $1$}};
%			\node at (1,3) [anchor=west] {{\scriptsize $4$}};
%			\node at (1,-0.05) [anchor=east] {{\scriptsize $q$}};
%			\draw(2,-1)--(2,-2)[dashed];
%			%			
%			\begin{scope}[xshift=4cm]
%				\draw(-2,-1)--(0,1)[];
%				\draw(0,1)--(-1,2)--(-1,3);
%				\draw(0,1)--(1,2)--(1,3);
%				\node at (-1,3) [anchor=west] {{\scriptsize $2$}};
%				\node at (1,3) [anchor=west] {{\scriptsize $3$}};
%				\node at (-1,0) [anchor=west] {{\scriptsize $q'$}};
%			\end{scope}
%		\end{tikzpicture}
\caption{
(a) Noise processes initially form isolated sets of pair-created anyons, 
each crossing the boundary of a tile. 
%These events are 
%each represented by curves, where anyons are denoted by specific 
%points on these curves. 
(b) To measure the total charge 
contained within each tile, 
we first join the participating curve 
diagrams arbitrarily into a single curve diagram.
%we must first join the fusion 
%trees associated with all such pairs. 
%This corresponds to joining 
%the edges associated with each pair into a continuous path, 
%representing a basis for the fusion space chosen arbitrarily. 
(c) The total charge of each tile can then be found 
by braiding anyons around each other until all charges within 
a tile are neighbors on the path, as in (d).  
The red lines correspond to the worldline braids that perform this step.
%(e) The specific braids required to transform (c) 
%to (d), where the labels of the four leaves of 
%these fusion trees correspond to the four anyons taken from 
%the top of the path in (c) to the bottom. 
%Dashed lines denote vacuum charges. 
%After applying several $R$ and $F$ moves (see Eqs.~(\ref{e:rmove}) and 
%(\ref{e:fmove})), the total charges within each tile can be read off the diagram as $q$ and $q'$.
}
\label{f:syndrome}
\end{center}
\vspace{-10pt}
\end{figure}

>>>>

We consider a 
2-dimensional system with
point-like excitations or quasiparticles known as anyons.
These anyons show exotic statistics under exchange.
The worldlines of these exchanges form \emph{braids}
in (2+1)-dimensions
which act unitarily on (the Hilbert space of) degrees of freedom known as the \emph{fusion space.}
To access this space we define
the \emph{observables} of the system: these
are regions bounded by circles (Wilson loops)
and measure the total charge of the region.
%The algebraic properties of these processes
These dynamics and observables obey algebraic rules
given by a unitary modular tensor category~\cite{Wang2010b}. % (UMTC).

@@

The state of a collection of anyons
can be described by the observables of the system.
An important difference between abelian and non-abelian
theories is this: the outcome of measurements in an abelian theory
are determined by particle content only.
That is, the fusion space is one-dimensional (a phase).
This is no longer the case in a non-abelian theory.
%An example is Fibonacci: $\tau\times\tau=\tau+\vac$
%\emph{picture here?}
%specify a state requires nesting multiple wilson lines.
In particular, we consider a system supporting Fibonacci
anyon excitations, denoted by $\tau.$
Two such anyons can have total charge
that is a superposition of $\tau$ and $\vac$ (vacuum),
so the fusion space in this case is 2-dimensional:
\begin{align*}
\includegraphics[]{pic-fibonacci.pdf}
\end{align*}
For $n$ anyons of type $\tau$, the 
dimension of this space grows asymptotically as $\varphi^n$ for $\varphi=\frac{1+\sqrt{5}}{2}$ 
the golden ratio. Additionally, there is a global degeneracy associated with the topology of the 
manifold on which the anyons reside. We will consider systems with the topology of a torus, 
which for the Fibonacci anyons gives rise to a 2-fold degeneracy.
%For convenience 
%(in particular to minimize finite-size effects in our numerical computations~\cite{Brell2013}), 
This global space is the one in which we will encode, and demonstrate protection of, quantum 
information.

Non-intersecting loops commute, and so
a basis for the state can be built from a maximal
set of disjoint, nested loops.
For three anyons, we relate two
possible ways of nesting loops %these measurements
via $F$-moves:
\begin{align}
\label{e:fmove}
\includegraphics[]{pic-fmove.pdf}
\end{align}
Here we have shown particles with charges $a$, $b$, and $c$
as well as total charge $d$.
The lhs shows a state where the anyons with charge $a$ and $b$ are
observed to have total charge $i$.
The rhs involves a superposition (summation)
over charge states $j$ of the anyons with charge $b$ and $c$.
The green line serves to linearly
order the anyons, and track
deformations in the underlying space 
(the history of braiding processes).
It also denotes the direction along
which $F$-moves occur.

%Here we show how braids effect observables:
%Here we show the $R$-move, corresponding to a (clockwise) braiding processes:
A clockwise braiding process is represented by an $R$-move:
\begin{align}
\label{e:rmove}
\includegraphics[]{pic-rmove.pdf}
\end{align}
%where $a$, $b$, and $c$ represent particle types, and $F$-moves,
%representing the associativity rule for fusion processes, 
This is meant to show a \emph{half-twist} of the space
inside the observable with value $c$
holding two anyons (punctures) with charges $a$ and $b$.
We are free to draw loops around charges,
and in general braiding will twist these loops in different
ways.
\dude{picture? ref Fig 2?}

These pictures, abstracted from the quantum context,
are called \emph{curve diagrams} and are used in
the study of braid groups \cite{Dehornoy2002}.
In general, such a curve diagram is 
a continuous embedding of a line that passes through 
each puncture once. 
Remarkably, any two such linear orderings are related by
braiding (half-twists).
\dude{optional paragraph..}

The $R$ and $F$-moves obey composition of diagrams.
\dude{picture?}
Fusion is the ``reverse'' of composition: replace the
interior of a disc by a single charge.
\dude{picture?}

For the Fibonacci anyons, the non-trivial $R$ and $F$ moves are 
\begin{equation*}
	R_{\vac}^{\tau\tau} = \e^{\frac{-4\pi i}{5}} 
	\ \ , \ \
	R_\tau^{\tau\tau}= \e^{\frac{3\pi i}{5}} 
	\ \ , \ \
	F_{\tau}^{\tau\tau\tau} = \begin{pmatrix}\varphi^{-1}&\varphi^{-\frac{1}{2}}\\\varphi^{-\frac{1}{2}}&-\varphi^{-1}\end{pmatrix} \,,
\end{equation*}
where the matrix is given in a basis labelled $(\vac,\tau)$.
%unitary actions on the fusion space. Pair-creation events should be considered isometries 
%that enlarge the fusion space, while fusion events correspond to projective measurements 
%of combined charge of some number of particles.
For detailed description of these 
processes in Fibonacci anyons, see e.g.~Ref.~\cite{Nayak2008} and references therein. 
Braiding of Fibonacci anyons and measurements of fusion outcomes is known to be 
sufficient to implement universal quantum computation~\cite{Freedman2002, Nayak2008}.

This treatment of anyon dynamics we consider to be a phenomenological model;
it neglects any microscopic details of the system.
This is consistent with the principles of topologically 
ordered systems and anyonic physics, where the key universal features describing the 
anyon model correspond to large length-scale physics, while the microscopic physics plays 
a less important (and non-universal) role.

@@

In order to keep track of the state of the system we maintain a set of disjoint curves, 
one for each connected component with trivial total charge.
The application of a pair-creation noise process corresponds to the 
addition of an extra piece of curve to the lattice supporting
two new anyons.
%along with two preferred points on it denoting the anyon locations. 
Following the noise processes, our simulation must %proceed to 
measure the total charge within each tile; the results of 
these measurements will form the error syndrome. 
This requires joining (arbitrarily) any curves that intersect that tile, 
and then performing $R$ and $F$-moves on the resulting
curve so that 
all anyons within the tile have been 
localized within a contiguous region of the curve. 
\dude{Heisenberg picture?}
The observable corresponding to the tile is then
represented in the state \dude{polish}.
%The total charge is then (probabilistically) read off the 
%relevant section of the fusion tree. 
An example of this 
procedure for a simple noise process is shown in \Fref{f:syndrome}.
%and the corresponding braid moves are shown in \Fref{f:syndrome}(e).
The combinatorial derivation of the algorithm and data structures used for 
this procedure will be discussed elsewhere.
%Braiding processes that encircle 
%a non-trivial loop of our (toric) manifold can also be 
%treated in an analogous way, following e.g.~\cite{Pfeifer2012}.
%\cggb{This paragraph and the next still needs another look.}

\begin{figure}[t!]
\begin{center}
	\includegraphics[width=1.0\columnwidth]{pic-syndrome.pdf}
\caption{
(a) Noise processes initially form isolated sets of pair-created anyons, 
each crossing the boundary of a tile. 
(b) To measure the total charge 
contained within each tile, 
we first join the participating curve 
diagrams arbitrarily into a single curve diagram.
(c) The total charge of each tile can then be found 
by braiding anyons around each other until all charges within 
a tile are neighbors on the curve, as in (d).  
The red lines correspond to the worldline braids that perform this step.
}
\label{f:syndrome}
\end{center}
\vspace{-10pt}
\end{figure}
